---
title: "cartography"
author: "Mia Forsline"
date: "9/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RColorBrewer)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(spData)
library(spDataLarge)
library(tidyverse)
library(tmap)

#GwR Ch.8 demo
library(raster)
library(dplyr)
library(leaflet)
```

8.2.1 tmap basics 
```{r}
# Add fill layer to nz shape
tm_shape(nz) +
  tm_fill() 
# Add border layer to nz shape
tm_shape(nz) +
  tm_borders() 
# Add fill and border layers to nz shape
tm_shape(nz) +
  tm_fill() +
  tm_borders() 
```

8.2.2 map objects 
```{r}
map_nz <- tm_shape(nz) + tm_polygons()
class(map_nz)
```

add a new shape using `tm_shape()`
```{r}
map_nz1 = map_nz +
  tm_shape(nz_elev) + tm_raster(alpha = 0.7)
map_nz1
```
add NZ territorial waters to the existing map object 
```{r}
nz_water = st_union(nz) %>% st_buffer(22200) %>% 
  st_cast(to = "LINESTRING")
map_nz2 = map_nz1 +
  tm_shape(nz_water) + tm_lines()
nz_water
map_nz2
```

Add NZ high points 
```{r}
map_nz3 = map_nz2 +
  tm_shape(nz_height) + tm_dots()
map_nz3
```
arrange multiple map objects into a single metaplot 
```{r}
tmap_arrange(map_nz1, map_nz2, map_nz3)
```

8.2.3 aesthetics 
```{r}
ma1 = tm_shape(nz) + tm_fill(col = "red")
ma2 = tm_shape(nz) + tm_fill(col = "red", alpha = 0.3)
ma3 = tm_shape(nz) + tm_borders(col = "blue")
ma4 = tm_shape(nz) + tm_borders(lwd = 3)
ma5 = tm_shape(nz) + tm_borders(lty = 2)
ma6 = tm_shape(nz) + tm_fill(col = "red", alpha = 0.3) +
  tm_borders(col = "blue", lwd = 3, lty = 2)
tmap_arrange(ma1, ma2, ma3, ma4, ma5, ma6)
```

```{r}
plot(st_geometry(nz), col = nz$Land_area) #base R functions accept nz$Land_area (a numeric vector)
tm_shape(nz) + tm_fill(col = "Land_area") #tmap aesthetic arguments want a character string 
```

set legend title 
```{r}
legend_title = expression("Area (km"^2*")")
map_nza = tm_shape(nz) +
  tm_fill(col = "Land_area", title = legend_title) + tm_borders()
map_nza
```

8.2.4 color settings
- `breaks` allows you to manually set the breaks
- `n` sets the number of bins into which numeric variables are categorized
- `palette` defines the color scheme, for example BuGn
```{r}
tm_shape(nz) + tm_polygons(col = "Median_income")
breaks = c(0, 3, 4, 5) * 10000
tm_shape(nz) + tm_polygons(col = "Median_income", breaks = breaks)
tm_shape(nz) + tm_polygons(col = "Median_income", n = 10)
tm_shape(nz) + tm_polygons(col = "Median_income", palette = "BuGn")
```

Breaks styles 
- style = "pretty", the default setting, rounds breaks into whole numbers where possible and spaces them evenly;
- style = "equal" divides input values into bins of equal range and is appropriate for variables with a uniform distribution (not recommended for variables with a skewed distribution as the resulting map may end-up having little color diversity);
- style = "quantile" ensures the same number of observations fall into each category (with the potential downside that bin ranges can vary widely);
- style = "jenks" identifies groups of similar values in the data and maximizes the differences between categories;
- style = "cont" (and "order") present a large number of colors over continuous color fields and are particularly suited for continuous rasters ("order" can help visualize skewed distributions);
- style = "cat" was designed to represent categorical values and assures that each category receives a unique color.