---
title: "ch.9 plotting spatial data"
author: "Mia Forsline"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(rnaturalearth)
library(tmap)
```

# 9.1 Every plot is a projection 
```{r}
w <- ne_countries(scale = "medium", returnclass = "sf")
plot(st_geometry(w))

st_is_longlat(w)
```

```{r}
DE = st_geometry(ne_countries(country = "germany", returnclass = "sf"))
DE.eqc = st_transform(DE, "+proj=eqc +lat_ts=51.14 +lon_0=90w")
#.eqc = "equidistant cylindrical" projection of PROJ
#lat_ts = projection parameter = latitude of true scale (one length unit N-S = one length unit E-W)

plot(DE)
plot(DE.eqc)
```

# 9.1.1 What is a good projection for my data ? 
The projection should preserve 
- areas 
- directions 
- some properties of distances 

# 9.2 Plotting points, lines, polygons, grid cells 

# 9.2.1 Colors 
- `RColorBrewer`
- `viridis`

# 9.2.2 Color breaks: `classInt` 

```{r}
library(classInt)
# set.seed(1) 
r = rnorm(100)
(cI <- classIntervals(r))
cI$brks
#stye can be: fixed, sd, equal, pretty, quantile, kmeans, hclust, bclust, fisher, or jenks
#style pretty may not obey n
#if n is missing, nclass.Sturges is used 
#if observations > 3,000 then a 10% sample is used to create breaks for fisher and jenks 
```

# 9.2.3 Graticule and other navigation aids 
- can show which projection was used 

# 9.3 Base `plot` 
- quick, usefule, exploratory plots 

# 9.3.1 Adding to plots with legends 

```{r}
library(sf)
nc = read_sf(system.file("gpkg/nc.gpkg", package = "sf"))
plot(nc["BIR74"], reset = FALSE, key.pos = 4)
#plot(st_buffer(nc[1,1], units::set_units(10, km)), col = 'NA', 
#     border = 'red', lwd = 2, add = TRUE)
```

```{r}
library(stars)
r = read_stars(system.file("tif/L7_ETMs.tif", package = "stars"))
circ = st_bbox(r) %>% st_as_sfc() %>% st_sample(5) %>% st_buffer(300)
hook = function() plot(circ, col = NA, border = 'yellow', add = TRUE)
plot(r, hook = hook, key.pos = 4)
```

# 9.3.2 Projections in base plots 
- base `plot` uses ellipsoidal coordinates using equirectangular projection using a latitude parameter equal to the middle latitude of the data bounding box 

# 9.4 Maps with `ggplot2` 

### tidy the data 
```{r}
suppressPackageStartupMessages(library(tidyverse))

nc.32119 = st_transform(nc, 32119) 
year_labels = c("SID74" = "1974 - 1978", "SID79" = "1979 - 1984")
nc_longer <- nc.32119 %>% 
  dplyr::select(SID74, SID79) %>% 
    pivot_longer(starts_with("SID")) 
```

### `geom_sf` developed in conjunction with `sf` 
```{r}
ggplot() + geom_sf(data = nc_longer, aes(fill = value)) + 
  facet_wrap(~ name, ncol = 1, labeller = labeller(name = year_labels)) +
  scale_y_continuous(breaks = 34:36) +
  scale_fill_gradientn(colours = sf.colors(20)) +
  theme(panel.grid.major = element_line(colour = "white"))
```
### `geom_stars` to go with `stars` package 
```{r}
library(ggplot2)
library(stars)
r = read_stars(system.file("tif/L7_ETMs.tif", package = "stars"))
ggplot() + geom_stars(data = r) +
        facet_wrap(~band) + coord_equal() +
        theme_void() +
        scale_x_discrete(expand = c(0,0)) + 
        scale_y_discrete(expand = c(0,0)) +
        scale_fill_viridis_c()
```


# 9.5 Maps with `tmap` 
```{r}
nc_longer <- nc.32119 %>% 
  dplyr::select(SID74, SID79) %>% 
    pivot_longer(starts_with("SID"), values_to = "SID") 

tm_shape(nc_longer) + tm_polygons("SID") + tm_facets(by = "name")
```
### `tmap` support for `stars` objects 
```{r}
tm_shape(r) + tm_raster()
```

# 9.6 Interactive maps: `leaflet`, `mapview`, `tmap`
```{r}
tmap_mode("view") 
#tmap mode set to interactive viewing 

tmap_mode("plot")
#tmap mode set to plotting (static map viewing)
```

# Ch. 5: Plotting simple features 

```{r}
qtm(nc)
```

```{r}
tmap_mode("view")
tm_shape(nc) + tm_fill("BIR74", palette = sf.colors(5))

ttm() #tmap mode set to plotting (for static maps)
tmap_last() #plot the last map we made, but now in the static view 

```

