---
title: "tmap tutorial"
author: "Mia Forsline"
date: "9/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RColorBrewer)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(spData)
library(spDataLarge)
library(tidyverse)
library(tmap)

#GwR Ch.8 demo
library(raster)
library(dplyr)
library(leaflet)
```

# 8.2.1 tmap basics 
```{r}
# Add fill layer to nz shape
tm_shape(nz) +
  tm_fill() 
# Add border layer to nz shape
tm_shape(nz) +
  tm_borders() 
# Add fill and border layers to nz shape
tm_shape(nz) +
  tm_fill() +
  tm_borders() 
```

8.2.2 map objects 
```{r}
map_nz <- tm_shape(nz) + tm_polygons()
class(map_nz)
```

add a new shape using `tm_shape()`
```{r}
map_nz1 = map_nz +
  tm_shape(nz_elev) + tm_raster(alpha = 0.7)
map_nz1
```
add NZ territorial waters to the existing map object 
```{r}
nz_water = st_union(nz) %>% st_buffer(22200) %>% 
  st_cast(to = "LINESTRING")
map_nz2 = map_nz1 +
  tm_shape(nz_water) + tm_lines()
nz_water
map_nz2
```

Add NZ high points 
```{r}
map_nz3 = map_nz2 +
  tm_shape(nz_height) + tm_dots()
map_nz3
```
arrange multiple map objects into a single metaplot 
```{r}
tmap_arrange(map_nz1, map_nz2, map_nz3)
```

# 8.2.3 aesthetics 
```{r}
ma1 = tm_shape(nz) + tm_fill(col = "red")
ma2 = tm_shape(nz) + tm_fill(col = "red", alpha = 0.3)
ma3 = tm_shape(nz) + tm_borders(col = "blue")
ma4 = tm_shape(nz) + tm_borders(lwd = 3)
ma5 = tm_shape(nz) + tm_borders(lty = 2)
ma6 = tm_shape(nz) + tm_fill(col = "red", alpha = 0.3) +
  tm_borders(col = "blue", lwd = 3, lty = 2)
tmap_arrange(ma1, ma2, ma3, ma4, ma5, ma6)
```

```{r}
plot(st_geometry(nz), col = nz$Land_area) #base R functions accept nz$Land_area (a numeric vector)
tm_shape(nz) + tm_fill(col = "Land_area") #tmap aesthetic arguments want a character string 
```

### set legend title 
```{r}
legend_title = expression("Area (km"^2*")")
map_nza = tm_shape(nz) +
  tm_fill(col = "Land_area", title = legend_title) + tm_borders()
map_nza
```

# 8.2.4 color settings
- `breaks` allows you to manually set the breaks
- `n` sets the number of bins into which numeric variables are categorized
- `palette` defines the color scheme, for example BuGn
```{r}
tm_shape(nz) + tm_polygons(col = "Median_income")
breaks = c(0, 3, 4, 5) * 10000
tm_shape(nz) + tm_polygons(col = "Median_income", breaks = breaks)
tm_shape(nz) + tm_polygons(col = "Median_income", n = 10)
tm_shape(nz) + tm_polygons(col = "Median_income", palette = "BuGn")
```

### Breaks styles 
- style = "pretty", the default setting, rounds breaks into whole numbers where possible and spaces them evenly;
- style = "equal" divides input values into bins of equal range and is appropriate for variables with a uniform distribution (not recommended for variables with a skewed distribution as the resulting map may end-up having little color diversity);
- style = "quantile" ensures the same number of observations fall into each category (with the potential downside that bin ranges can vary widely);
- style = "jenks" identifies groups of similar values in the data and maximizes the differences between categories;
- style = "cont" (and "order") present a large number of colors over continuous color fields and are particularly suited for continuous rasters ("order" can help visualize skewed distributions);
- style = "cat" was designed to represent categorical values and assures that each category receives a unique color.

`palette` argument expects a vector of colors or a new color palette name, which can be selected interactively with `tmaptools::palette_explorer()`

### Color Palettes:
1. categorical
2. sequential (singles such as `Blues` or multi-color such as `YlOrBr`)
```{r}
tm_shape(nz) + tm_polygons("Population", palette = "Blues")
tm_shape(nz) + tm_polygons("Population", palette = "YlOrBr")
```

3. Diverging = visualize the difference from an important reference point, which can be adjusted using `midpoint` argument 

# 8.2.5 Layouts 
Include objects to be mapped, title, scale bar, margins, aspect ratios
```{r}
map_nz + 
  tm_compass(type = "8star", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200), text.size = 1)

map_nz + tm_layout(title = "New Zealand")
map_nz + tm_layout(scale = 5)
map_nz + tm_layout(bg.color = "lightblue") #background color 
map_nz + tm_layout(frame = FALSE) #remove background box outline
```

### Layout styles 
```{r}
map_nza + tm_style("bw")
map_nza + tm_style("classic")
map_nza + tm_style("cobalt")
map_nza + tm_style("col_blind")

#tmap_style_catalogue()
```

# 8.2.6 Faceted maps 
```{r}
urb_1970_2030 = urban_agglomerations %>% 
  filter(year %in% c(1970, 1990, 2010, 2030))

tm_shape(world) +
  tm_polygons() +
  tm_shape(urb_1970_2030) +
  tm_symbols(col = "black", border.col = "white", size = "population_millions") +
  tm_facets(by = "year", nrow = 2, ncol = 2, free.coords = FALSE)
#free.coords = each map has its own bounding box 
```

# 8.2.7 Inset maps 
```{r}
#define the area of interest 
nz_region = st_bbox(c(xmin = 1340000, xmax = 1450000,
                      ymin = 5130000, ymax = 5210000),
                    crs = st_crs(nz_height)) %>% 
                    st_as_sfc()

#create the map 
nz_height_map = tm_shape(nz_elev, bbox = nz_region) +
  tm_raster(style = "cont", palette = "YlGn", legend.show = TRUE) +
  tm_shape(nz_height) + tm_symbols(shape = 2, col = "red", size = 1) +
  tm_scale_bar(position = c("left", "bottom"))
nz_height_map

#create the inset map 
nz_map = tm_shape(nz) + tm_polygons() +
  tm_shape(nz_height) + tm_symbols(shape = 2, col = "red", size = 0.1) + 
  tm_shape(nz_region) + tm_borders(lwd = 3) 
nz_map 

#arrange inset map onto main map 
library(grid)
nz_height_map
print(nz_map, vp = viewport(0.8, 0.27, width = 0.5, height = 0.5))
```

### map of the contiguous US
```{r}
us_states_map = tm_shape(us_states, projection = 2163) + tm_polygons() + 
  tm_layout(frame = FALSE)
us_states_map
```

### Hawaii + Alaska Maps 
```{r}
hawaii_map = tm_shape(hawaii) + tm_polygons() + 
  tm_layout(title = "Hawaii", frame = FALSE, bg.color = NA, 
            title.position = c("LEFT", "BOTTOM"))
alaska_map = tm_shape(alaska) + tm_polygons() + 
  tm_layout(title = "Alaska", frame = FALSE, bg.color = NA)
hawaii_map
alaska_map
```
### Arrange all 3 maps 
```{r}
us_states_map
print(hawaii_map, vp = grid::viewport(0.35, 0.1, width = 0.2, height = 0.1))
print(alaska_map, vp = grid::viewport(0.15, 0.15, width = 0.3, height = 0.3))
```

# 8.3 Animated maps 
```{r}
#create the frames 
#note: along = "year" rather than by = "year"
urb_anim = tm_shape(world) + tm_polygons() + 
  tm_shape(urban_agglomerations) + tm_dots(size = "population_millions") +
  tm_facets(along = "year", free.coords = FALSE)
urb_anim
```

```{r}
#create and save the gif file 
tmap_animation(urb_anim, filename = "urb_anim.gif", delay = 25)
```

# 8.4 Interactive Maps 
```{r}

```

